---
title: "RF"
author: "Kylian van Noordenne"
date: "22-11-2020"
output: html_document
---

Within this section, a random forest will be created. FIrst a preprocessing recipe is created. 

```{r libraries }
library(ranger)
library(doParallel)
library(themis)
```

```{r Recipe}
rf_recipe <- recipe(price ~ bedrooms + beds + host_identity_verified + guests_included + extra_people + pool, data = data_train) 
rf_recipe
```

Specify random forest with 500 trees
```{r }
tune_spec <- rand_forest(mtry = tune(), trees = 500) %>%
  set_engine("ranger")
```

Combine the recipe and the model into a workflow that can be tuned
```{r }
tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(tune_spec)
```

We create a metric set that calculates the Root Mean Square Error (rmse), the Mean Absolute Error (mae) and the R-squared (rsq_trad)
```{r }
class_metrics <- metric_set(rmse, mae, rsq_trad)
```

The command bellow allows us to do computations in paallel
```{r }
registerDoParallel()
```

```{r }
set.seed(12345)
tune_res <- tune_grid(
  tune_wf,
  resamples = data_folds,
  grid = 10,
  metrics = class_metrics
)
tune_res
```

```{r }
tune_res %>%
  collect_metrics()
```
A plot for finding the best mtry, based on the criteria of the rmse. A lower rmse would indicate a better results, as a lower value indicates a lower error of prediction. 
```{r }
tune_res %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>% 
  ggplot(aes(x = mtry, y = mean)) +
  geom_line() +
  geom_point() +
  labs(y = "rmse")
```

!!! CHANGE !!!! this command allows me to see 
```{r }
tune_res %>% show_best("rmse")
```
Now select the best model based on the rmse criteria
```{r }
best_rmse <- select_best(tune_res, "rmse")
final_rf <- finalize_workflow(tune_wf, best_rmse)
final_rf
```

Now we can train the finalized workflow on our entire training rest
```{r }
final_res <- final_rf %>%
  last_fit(data_split, metrics = class_metrics)
```

The results based on the test set will be 
```{r }
final_res %>%
  collect_metrics()
```

```{r }

```

```{r }

```